HyperLogLog: Exponentially Bad in Adversarial Settings

### 2020-05-05 docs/HLL_exp_bad/main.tex
draft S1

\begin{itemize}
\item[S1:] The adversary has access to an empty black-box replica of the sketch in use. He can use the API to insert elements into that shadow device, get the cardinality estimate, and reset the sketch. The attacker does not know the elements previously added by honest users nor the state. That is we can test and find elements that are unlikely to increase the HLL estimate on the shadow HLL and then insert the on the target HLL but we do not know the state of the target HLL. It would also be reasonable for the provider to transform the items/hash with  a salt e.g. with a different salt per customer. So one can have both an open source implementation and not know the hash.
\item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''. In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack. This kind of attack scenario is appropriate in the context of the port scanning attack of~\cite{portscanhll}.
\item[S3:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use. Additionally, it can access and interact with the sketch via an API provided by the sketch owner, allowing it to insert items into the HLL sketch and ask for the HLL's cardinality estimate at any point in time.
\item[S4:] The adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch \emph{at one specific point in time}, i.e.\ it is given a \emph{snapshot} of the sketch. It then causes chosen inputs to be inserted into the HLL sketch. An S4 adversary can compute a lower bound on the HLL cardinality estimate based on its snapshot and whatever it causes to be inserted into the HLL sketch during its attack (this is irrespective of whether its attack proceeds with input from other users being added to the HLL sketch in parallel with its attack or not; if it knows that there is no other input, then it can compute the \emph{exact} value of the HLL cardinality estimate at every point in its attack instead of a lower bound).
\end{itemize}

@@

We start by highlighting the fact that S1 seems unrealistic in practice. In a networking application, the authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the attacker learns which machine is used to do the monitoring and buys the same device. It is reasonable to think that the provider may use a different salt per device, which would make any effect on the shadow device unexpected on the target and would break the whole attack presented in Section 3.2 of \cite{hllvuln}. We consider that the machines use the same exact internals. Assuming that such setup is possible, we show that it is actually possible for the attacker to learn the internals of the HLL sketch. The API must explicitly tell the customer the error of the sketch, and hence the number of buckets. We show that the attacker can learn the hash of a targeted item using the knowlwdege of S1, and thus end up in the S2 scenario. To learn $h(x)$, the attacker can insert the item $x$ on en empty sketch, and ask for the resulting cardinality, which we denote $c$. By solving:
\begin{eqnarray*}
    \alpha 2^{2n}(\sum_{j=1}^{2^n}2^{-c_j})^{-1}& = & c \\
    \Leftrightarrow log_2(c)-log_2(\alpha m^2 - cm + c) &=& M[i].
\end{eqnarray*}
where  the attacker can learn the value of the updated estimator of the bucket $x$ was mapped to, and recover some bits of $h(x)$. This may be possible under certain circumstances, but we do not consider it typical of HLL deployments, and furthermore, it is easy for the attacker to escalate to the knowledge of S2.

>>>>

\begin{itemize}
\item[S1:] The adversary has access to an empty black-box replica of the sketch in use. He can use the API to insert elements into that shadow device, get the cardinality estimate, and reset the sketch. The attacker does not know the state of the targeted HLL sketch, i.e., the cardinality estimation nor elements previously added by other users.
\item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''. In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack. This kind of attack scenario is appropriate in the context of the port scanning attack of~\cite{portscanhll}.
\item[S3:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use. Additionally, it can access and interact with the sketch via an API provided by the sketch owner, allowing it to insert items into the HLL sketch and ask for the HLL's cardinality estimate at any point in time.
\item[S4:] The adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch \emph{at one specific point in time}, i.e.\ it is given a \emph{snapshot} of the sketch. It then causes chosen inputs to be inserted into the HLL sketch. An S4 adversary can compute a lower bound on the HLL cardinality estimate based on its snapshot and whatever it causes to be inserted into the HLL sketch during its attack (this is irrespective of whether its attack proceeds with input from other users being added to the HLL sketch in parallel with its attack or not; if it knows that there is no other input, then it can compute the \emph{exact} value of the HLL cardinality estimate at every point in its attack instead of a lower bound).
\end{itemize}

@@

We start by highlighting the fact that S1 seems unrealistic in practice. In a networking application, the authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the attacker learns which machine is used to do the monitoring and buys the same device. It is possible that the provider uses a different salt/key per device, which would make any pre-computation irrelevant, and considerably weaken the attack presented in Section 3.2 of \cite{hllvuln}. If we consider that the devices are perfect copies, we show that it is actually possible for the attacker to learn the internals of the HLL sketch. Precisely, the attacker can recover some bits of interest in the hash of targeted items, and infer the knowledge given to the adversary under S2. To do so, the attacker can insert the item $x$ on an empty sketch, ask for the resulting cardinality, which we denote $c$, and solve:
\begin{eqnarray*}
    \alpha 2^{2n}(\sum_{j=1}^{2^n}2^{-c_j})^{-1}& = & c \\
    \Leftrightarrow log_2(c)-log_2(\alpha 2^{2n} - c2^n + c) &=& c_i
\end{eqnarray*}
to learn $c_i$. $\alpha$ is a constant, $2^n$ is known as the provider must explicitly tell the customer the expected error of the estimation, hence the number of buckets, and $i$ is an arbitrary value representing the index of the bucket $x$ is mapped to (recovering $i$ is not useful). As $c_i$ denotes the $i$ bucket's content, updated by the insertion of $x$, $c_i-1$ is (an approximation of) the number of leading zeros in the $l$ rightmost bits of $h(x)$, which is the information extracted from $h(x)$ and used by the adversary in his attack under S2, as we detail in Section~\ref{sec:attacks}. %the value of the updated estimator of the bucket $x$ was mapped to, and recover some bits of $h(x)$.
To summarize, we do not consider this scenario typical of HLL deployments, and even if such setup may be possible under certain circumstances, the attacker can infer the knowledge given in S2.

\mathilde{Do you think that we should mention the idea behind the attack of \cite{hllvuln} Section 3.2 ? As they explained in their emails "The attacker can then test and find elements that do not increase the HLL estimate on the shadow HLL and thus will unlikely increase the cardinality of the target HLL when inserted."}
